Large Scale Structure Reconstruction with Short-Wavelength Modes

### 2019-10-22 paper/v2.tex
1

\end{abstract}
\maketitle
\noindent {\it Introduction.}
Measuring the distribution of matter on large scales is one of the goals of cosmological surveys. The information contained on large scales may provide information about issues ranging from the turnover in the power spectrum (and therefore the total matter density) to the accelerated universe to anomalies observed in the cosmic microwave background to primordial nongaussianity. %\peikai{, through which we can have a better constraint of $f_{\rm NL}$ or solve large scale anomalies,} despite it is hard to accomplish directly.\RC{Should include brief examples of why it is very interesting - for example to constrain fNL, large scale anomalies, detect matter radiation-equality turnover - make it clear that a measurement would be exciting} 
Direct measurements are difficult because of systematic effects, so indirect approaches have been considered. Small scale structure grows differently in the presence of an large-scale overdensity: it is as if the mean background density is larger than on average. This relation between long- and short-wavelength modes has been discussed in recent years \cite{Baldauf:2011fer}\cite{Barreira:2017res}. The method of using the small scale position-dependent power spectrum to compute the squeezed-limit bispectrum also indicates that small scale perturbations can be used to infer large scale information \cite{Chiang:2014pos}\cite{Chiang:2015poss}. In order to fully achieve this goal, here we construct a quadratic estimator to measure long-wavelength modes indirectly.

@@

We begin with a brief review of SPT up to second-order, build the quadratic estimator, and then calculate its detectability. We then apply the estimator to data from a large N-body simulation and demonstrate that it successfully extracts the large scale modes. %Finally we conclude this work and discuss future work related to this in section \ref{sec4}. 
We use a flat $\Lambda$CDM model with Planck Collaboration XVI (2014)~\cite{Planck:2014cos} cosmological parameters in this work (to match the parameters of the N-body simulation).

\noindent {\it Standard Perturbation Theory.}\label{sec1}
%In this section we review the SPT approach which describes small scale nonlinearity. \RC{Say something about: see also a textbook, or the Bernardeau review reference - it looks to me that some of your notation is from the latter. Stating this explicitly could be helpful for a reader.} \peikai{See also \cite{Bernardeau:2002rev} for a detailed derivation of SPT.} We then compute the off-diagonal term of two short-wavelength modes in the squeezed limit. This non-zero off-diagonal term proportional to a long-wavelength mode will be the key factor of constructing the quadratic estimator. \\
Starting from a perfect pressureless fluid, the nonrelativistic cosmological fluid equations are the continuity, Euler and Poisson equations:
\bea
\frac{\partial \delta(\vec{x},\tau)}{\partial \tau} &+&\vec{\nabla}\times [(1+\delta(\vec{x},\tau))\vec{v}(\vec{x},\tau)] =0\eql{fluid1} \\
(\frac{\partial}{\partial \tau} +\vec{v}(\vec{x},\tau)&\times&\vec{\nabla})\vec{v}(\vec{x},\tau)=-\frac{da}{d\tau}\frac{\vec{v}(\vec{x},\tau)}{a}-\vec{\nabla}\Phi \eql{eulereq}\\
&&\nabla^2 \Phi = 4\pi G a^2 \bar{\rho}_{\rm m} \delta(\vec{x},\tau) .\eql{fluid3}
\eea
These equations fully determine the time evolution of the local density contrast $\delta$ and the peculiar velocity field $\vec{v}=d\vec{x}/d\tau$.
We can solve these equations perturbatively in Fourier space~\cite{Bernardeau:2002rev}:
\bea
{\delta}(\vk,\tau) &=&\sum_{n=1}^{\infty} {\delta}^{(n)}(\vk,\tau)=\sum_{n=1}^{\infty}D_1^{n}(\tau)\delta_{n}(\vk)  \eql{pert1}\\
{\theta}(\vk,\tau)&=&\sum_{n=1}^{\infty}{\theta}^{(n)}(\vk,\tau) \vs
&=&-\frac{d\,\ln D_1(\tau)}{d\tau}\sum_{n=1}^{\infty}D_1^{n}(\tau)\theta_{n}(\vk) \eql{pert2}
\eea
where $D_1$ is the linear growth factor. %\RC{define this e.g. $\delta(\tau)=D_{1}\delta(0)$} \peikai{defined by $\delta^{(1)}(\vk,\tau)=D_{1}(\tau)\delta_{0}(\vk)$ with $\delta_0$ being the initial density field. Is it necessary? This equation is already indicated in Eq(4).}. 
The first order term ${\delta}^{(1)}$ corresponds to linear evolution. The linear power spectrum is given by this first order term via:
\be 
\langle {\delta^{(1)}}(\vk,\tau){\delta^{(1)}}(\vk',\tau) \rangle =(2\pi)^3 \delta_{\rm D}(\vk+\vk')P_{\rm lin}(k,\tau) \eql{lin}
\ee 
Substituting the perturbative series \ec{pert1} and \ec{pert2} into the Fourier transformed fluid equations \ec{fluid1}-\ec{fluid3} leads to an expression for the second-order density contrast:
\be
{\delta}^{(2)}(\vk,\tau)=\int \frac{d^{3}\vk_{1}}{(2\pi)^3} F_2(\vk_1,\vk-\vk_1){\delta}^{(1)}(\vk_1,\tau) {\delta}^{(1)}(\vk-\vk_1,\tau) \eql{sorder} 
\ee
with 
\be
F_{2}(\vk_1,\vk_2)=\frac{5}{7}+\frac{2}{7}\frac{(\vk_1\cdot \vk_2)^2}{k_1^2 k_2^2}+\frac{\vk_1\cdot \vk_2}{2k_1k_2}(\frac{k_1}{k_2}+\frac{k_2}{k_1}).\eql{f2}
\ee
Note that \ec{pert1}, \ec{pert2} and \ec{f2} are completely accurate only in an Einstein-de Sitter universe. Nonetheless for related calculations in a $\Lambda$CDM universe, the difference is found to be negligible \cite{Takahashi:2008to}, and thus we use the expressions from \ec{pert1} and \ec{f2} throughout this work.\\

Using this expression for $\delta^{(2)}$, we can calculate the two-point correlation of two short-wavelength modes $\vk_s$ and $\vk_s'$, in the squeezed limit $\vk_l=\vk_s+\vk_s'$ with $\vk_s,\vk_s' \gg \vk_l$. Here $\vk_l$ corresponds to a long-wavelength mode, and we suppress the time dependence. To second order,
\bea 
 \langle {\delta}(\vec{k}_s){\delta}(\vec{k}_s') \rangle|_{\vk_s+\vk_s'=\vk_l}&=&
  \langle {\delta}^{(1)}(\vec{k}_s){\delta}^{(2)}(\vec{k}_s') \rangle\vs
  &+&\langle {\delta}^{(2)}(\vec{k}_s){\delta}^{(1)}(\vec{k}_s') \rangle.
\eea 
Substituting \ec{sorder} into the first bracket we get:
\bea 
\langle {\delta}^{(1)}(\vec{k}_s){\delta}^{(2)}(\vec{k}_s') \rangle =  \int \frac{d^3\vec{k}}{(2\pi)^3} F_2 (\vec{k},\vec{k}_s'-\vec{k})\vs
\times \langle {\delta}^{(1)}(\vec{k}_s){\delta}^{(1)}(\vec{k}_s'-\vec{k}){\delta}^{(1)}(\vec{k}) \rangle \eql{integral}
\eea 
%\peikai{Need to rephrase (don't know how): We claim that generally correlation function of three Gaussian fields is non-zero under following consideration: if a three-point correlation function consists of two short-wavelength modes and one long-wavelength mode, we can do the following contraction:
%\be 
%\langle
%\contraction{}{{\delta}}{{\delta^{(1)}\;\;\;\;}(}{}
%{\delta^{(1)}}(\vk_{1}) {\delta^{(1)}}(\vk_{2}){\delta^{(1)}}(\vk_3) 
%\rangle
%=\langle {\delta^{(1)}}(\vk_{1}) {\delta^{(1)}}(\vk_{2}) \rangle {\delta^{(1)}}(\vk_3)  \eql{contraction}
%\ee 
%$\vk_1$ and $\vk_2$ corresponds to two arbitrary short modes and $\vk_3$ arbitrary long mode. The above equation is true since in a real life survey, we can only measure this long-wavelength mode very limited times. Thus it won't have any statistical property and we can safely extract the long-wavelength mode out of the bracket.}\\
When one of the wavenumbers in the 3-point function in \ec{integral} is very small, that mode can be considered as a background mode. The small scale modes evolve in the presence of whatever long wavelength modes happen to be present. Therefore, 
\bea
\langle
%\contraction{}{{\delta}}{{\delta^{(1)}\;\;\;\;}(}{}
\delta^{(1)}(\vk_{s}) \delta^{(1)}(\vk_{s}-\vk){\delta^{(1)}}(\vk) 
\rangle
&=&\langle {\delta^{(1)}}(\vk_{s}) {\delta^{(1)}}(\vk_{s}-\vk) \rangle {\delta^{(1)}}(\vk)  \vs
&+&\langle {\delta^{(1)}}(\vk_{s}) {\delta^{(1)}}(\vk) \rangle {\delta^{(1)}}(\vk_s'-\vk) .
\vs
\eql{contraction}
\eea
The first term on the right occurs when $\vk$ is small and the second when $\vk_s'-\vk$ is small.

%\RC{Maybe this?: The RHS of \ec{integral} includes an integral over the three point correlation function of modes of a Gaussian field. If one of the modes has a much longer wavelength than the others, then its contribution will be limited to a few examples, and a statistical average will not take place.  In this case, we can move the long wavelength mode outside the bracket, so that the following is a good approximation:
%\be 
%\langle
%\contraction{}{{\delta}}{{\delta^{(1)}\;\;\;\;}(}{}
%{\delta^{(1)}}(\vk_{1}) {\delta^{(1)}}(\vk_{2}){\delta^{(1)}}(\vk_3) 
%\rangle
%=\langle {\delta^{(1)}}(\vk_{1}) {\delta^{(1)}}(\vk_{2}) \rangle {\delta^{(1)}}(\vk_3)  \eql{contraction},
%\ee 
%where $\vk_1$ and $\vk_2$ are two short modes and $\vk_3$ is the long mode.}\\
%We can use this contraction twice in \ec{integral}. One use occurs when $|\vk| \ll |\vk_s|,\, |\vk_s'-\vk|$, when we can extract ${\delta}^{(1)}(\vk)$; the other is when $|\vk_s'-\vk| \ll |\vk|,\,|\vk_s|$.\\
Using \ec{lin}, \ec{integral} then becomes:
\bea 
&&\int \frac{d^3\vec{k}}{(2\pi)^3} F_2 (\vec{k},\vec{k}_s'-\vec{k})\langle {\delta}^{(1)}(\vec{k}_s){\delta}^{(1)}(\vec{k}_s'-\vec{k}){\delta}^{(1)}(\vec{k}) \rangle \vs
&=&\int d^3\vk F_2(\vec{k},\vec{k}_s'-\vec{k}) \delta_{\rm D}(\vk_s+\vk_s'-\vk)P_{\rm lin}(k_s){\delta}^{(1)}(\vec{k}) \vs
&+&\int d^3\vk F_2(\vec{k},\vec{k}_s'-\vec{k}) \delta_{\rm D}(\vk_s+\vk)P_{\rm lin}(k_s){\delta}^{(1)}(\vk_s'-\vk) \vs
&=&2F_2(-\vk_s,\vk_s+\vk_s')P_{\rm lin}(k_s){\delta}^{(1)}(\vk_s+\vk_s')
\eea 
where we take advantage of the fact that $F_2$ is a symmetric function. Finally we have:
\be 
\langle {\delta}(\vec{k}_s){\delta}(\vec{k}_s') \rangle =f(\vec{k}_s,\vec{k}_s'){\delta}^{(1)}(\vec{k}_l) \eql{2pt}
\ee 
with
\bea
f(\vec{k}_s,\vec{k}_s')&=&2F_2(-\vec{k}_s,\vec{k}_s+\vec{k}_s')P_{\rm lin}(k_s)\vs
&+&2F_2(-\vec{k}_s',\vec{k}_s+\vec{k}_s')P_{\rm lin}(k_s')       
\eea 
\ec{2pt} suggests that we can estimate long-wavelength modes using short-wavelength modes.

\noindent{\it Quadratic Estimator.}\label{sec2}
We can now construct the quadratic estimator for long-wavelength modes starting from \ec{2pt} and summing over as many pairs as possible with weights that maximize the signal to noise. As with the case of CMB lensing, we can write the general form of the estimator by averaging over pairs of short-wavelength modes:
\begin{eqnarray}
\hat{\delta}^{(1)}(\vec{k}_l)=A(\vec{k}_l)\int \frac{d^3 \vec{k}_s}{(2\pi)^3} g(\vec{k}_s,\vec{k}_s'){\delta}(\vec{k}_s){\delta}(\vec{k}_s') \eql{quadest},
\end{eqnarray} 
with $g$ being a weighting function, $\vk_s'=\vk_l-\vk_s$ and $A$  defined by requiring that $\langle \hat{\delta}^{(1)}(\vec{k}_l) \rangle={\delta}^{(1)}(\vec{k}_l)$:
\begin{eqnarray}
A(\vec{k}_l)=\bigg[\int \frac{d^3 \vec{k}_s}{(2\pi)^3} g(\vec{k}_s,\vec{k}_s')f(\vec{k}_s,\vec{k}_s')  \bigg]^{-1} \eql{a}
\end{eqnarray}
In the absence of shot noise the Gaussian noise is given by:
\be 
\langle \hat{\delta}^{(1)}(\vk_{l})\hat{\delta}^{(1)}(\vk_{l}') \rangle = (2\pi)^3 \delta_{\rm D}(\vk_{l}-\vk_{l}')[P_{\rm lin}(k_{l})+N(\vk_l)]
\ee 
with 
\begin{eqnarray}
&&N(\vec{k}_{l})=2A^2(\vk_{l})\vs
&&\times\int \frac{d^3 \vec{k}_{s}}{(2\pi)^3} g^2(\vec{k}_{s},\vk_l-\vec{k}_{s})P_{\rm nl}(k_{s})P_{\rm nl}(|\vk_l-\vk_s|)
\end{eqnarray}
where $P_{\rm nl}$ is the nonlinear power spectrum. Minimizing the noise term we can fix the form of $g$ to be:
\begin{eqnarray}
g(\vec{k}_{s},\vec{k}_{s}')=\frac{f(\vec{k}_{s},\vec{k}_{s}')}{2P_{\rm nl}(k_{s})P_{\rm nl}(k_{s}')}
\end{eqnarray} 
The noise term reduces simply to $N(\vk_l)=A(\vk_l)$. We find by testing that the value of $N$ is very insensitive to the choice of the lower limit of the integration \ec{a}, since most of the contribution comes from large $k_s$. %We discuss the choice of the upper limit later, in section \ref{sec3}.\\

Assuming Gaussian noise, the projected detectability of a $P(k_l)$ measurement using the quadratic estimator can be expressed as:
\be
\frac{1}{\sigma^{2}(k_l)}=\frac{V k_l^2 \Delta k }{(2\pi)^2}\bigg[\frac{P_{\rm lin}(k_l)}{P_{\rm lin}(k_l)+N(k_l) }\bigg]^2 \eql{error},
\ee
where $V$ is the volume of a survey and we compute the detectability for a set of narrow $k_l$-bins each separated by width $\Delta k$.
%\RC{ Write a couple of sentences about Figure 1, bearing in mind that you have put information in the caption already. Also remember to change the survey volume to match DESI and state this.} 
In Fig.~\rf{SN}, we show the projected errors on the long-wavelength power spectrum using this quadratic estimator in a large survey. The upper limit of the $\vk_s$ integration in \ec{a} is set to be $0.22\,h\,\rm Mpc^{-1}$. We will see that this choice of the upper limit is reasonable for our current construction at $z=0$. Also notice that $P_{\rm lin}(k_l)$ dominates over $N(k_l)$ in \ec{error} for this upper limit, thus the projected error bars are only slightly wider than the cosmic variance error bars ($N=0$).

%\peikai{plot linear or nonlinear? Change ylabel?}
\Sfig{SN}{Long-wavelength power spectrum and its error from \ec{error} which can be expressed as $P(k_l)\sigma(k_l)$. We assume a toy survey of boxsize $L=2.5\, h^{-1}\rm\, Gpc$, thus volume $V=L^3$ and width $\Delta k = 2\pi/L$. Moreover, we set the integration range for $\vk_s$ from $0.03 \,h \,\rm Mpc^{-1}$ to $0.22\,h\,\rm Mpc^{-1}$.}

@@

We use the code nbodykit \cite{Hand:2018nby} to measure the Fourier density modes, and \ec{quadest} to estimate the long wavelength modes  from the measured short wavelength modes. How well the estimator works can be seen from Fig.~\rf{hist}, where we show histograms of the ratio of the estimated mode amplitudes $\hat{\delta}(\vk_l)$ to their true amplitudes $\delta(\vk_l)$ for different values of $\vk_l$. The two panels show the differences between a short wavelength mode cutoff of $k_s=0.22\, h\, \rm Mpc^{-1}$ and  $k_s=0.37\,h\, \rm Mpc^{-1}$. Notice that second-order SPT becomes less accurate as shorter wavelengths are used and will produce a bias of our quadratic estimator. The figure shows that individual mode amplitudes are  unbiased when $k_{s,\rm max}= 0.22\,h\, \rm Mpc^{-1}$, while for $k_{s,\rm max}=0.37\,h\, \rm Mpc^{-1}$, the results are biased (the center of the ratio is $\sim 20 \%$ too high). %and more incorrect (about $\sim 50 \%$ of the modes have negative real parts of $\delta/\hat{\delta}$) due to the increasing inaccuracy of second-order SPT. } 
%\RC{Comment on what the figure shows -  for example individual mode amplitudes are unbiased when $k_s= 0.15 \rm Mpc^{1}$, and predicted to within $\sim x \%$} \RC{moved this here from fig caption: We can see that the results for $k_s=0.25 \rm Mpc^{1}$ are biased (the mode is $\sim 20 \%$ too high) due to the increasing inaccuracy of second-order SPT as we move into nonlinear scales}
\Sfig{hist}{Comparison of measured and predicted Fourier mode amplitudes: histograms of number counts of $|\hat{\delta}(\vk_l)/\delta(\vk_l)|$ for two different $k_s$ integration ranges.}

@@

% of this approach is to reproduce large scale structure in real space. 
%To test this, we compute real space density fields by taking the inverse transform of both the directly measured long wavelength Fourier modes and the long wavelength modes predicted using the quadratic estimator. 
%These density fields, $\delta$ and $\hat{\delta}$ are plotted in . %We can see that \peikai{our quadratic estimator }
%\RC{describe how well it works based on what we can see in the figure}
The bottom panel shows the difference between the estimated and true density fields. It is apparent that the differences are much smaller than the overdensities; equivalently the estimator does an excellent job of extracting the large scale density field.

\noindent{\it Conclusion.}\label{sec4}
In this paper, we have proposed a new and potentially powerful method to measure long-wavelength modes without having to actually measure large scale structure directly. The estimator works well on an N-Body simulation, so applying this estimator to survey data is an exciting next step. Among the issues that must be faced when dealing with a spectroscopic galaxy survey are: galaxy bias, redshift space distortions \cite{Kaiser:1987rsd}, and light-cone effects. We do not expect any of these to be show-stoppers, so it is tempting to speculate about the possibilities that will open up with this estimate of the large scale density field. 
First, we can hope to measure 3D clustering on scales larger than the matter radiation equality turnover without worrying about large-scale systematic effects. General relativistic effects are strongest on large scales (e.g., \cite{Jeong:2012ls}), and these could be detected. There is evidence of large scale anomalies, in the CMB, that could be confronted with maps of large scale structure obtained with this estimator. Primordial non-Gaussianity generated by inflation leaves an imprint on the largest scales. There is even the possibility of cross-correlating the large-scale matter field with the CMB itself to extract information about the longest wavelength modes in the universe. Since the current epoch of acceleration is a large-scale, late-time effect, there is the possibility of learning about the mechanism responsible for acceleration. Although, as mentioned above, challenges remain, there is also the possibility of using even smaller wavelength modes in our estimator by going to higher order in perturbation theory.

>>>>

\end{abstract}
\maketitle
\noindent {\it Introduction.}
Measuring the distribution of matter on large scales is one of the goals of cosmological surveys. The information contained on large scales may provide information about issues ranging from the turnover in the power spectrum (and therefore the total matter density) to the accelerated universe to anomalies observed in the cosmic microwave background to primordial nongaussianity. Direct measurements are difficult because of systematic effects, so indirect approaches have been considered. Small scale structure grows differently in the presence of an large-scale overdensity: it is as if the mean background density is larger than on average. This relation between long- and short-wavelength modes has been discussed in recent years \cite{Baldauf:2011fer}\cite{Barreira:2017res}. The method of using the small scale position-dependent power spectrum to compute the squeezed-limit bispectrum also indicates that small scale perturbations can be used to infer large scale information \cite{Chiang:2014pos}\cite{Chiang:2015poss}. In order to fully achieve this goal, here we construct a quadratic estimator to measure long-wavelength modes indirectly.

@@

We begin with a brief review of SPT up to second-order, build the quadratic estimator, and then calculate its detectability. We then apply the estimator to data from a large N-body simulation and demonstrate that it successfully extracts the large scale modes. 
We use a flat $\Lambda$CDM model with Planck Collaboration XVI (2014)~\cite{Planck:2014cos} cosmological parameters in this work (to match the parameters of the N-body simulation).\\

\noindent {\it Standard Perturbation Theory.}\label{sec1}
Starting from a perfect pressureless fluid, the nonrelativistic cosmological fluid equations are the continuity, Euler and Poisson equations:
\bea
\frac{\partial \delta(\vec{x},\tau)}{\partial \tau} &+&\vec{\nabla}\times [(1+\delta(\vec{x},\tau))\vec{v}(\vec{x},\tau)] =0\eql{fluid1} \\
(\frac{\partial}{\partial \tau} +\vec{v}(\vec{x},\tau)&\times&\vec{\nabla})\vec{v}(\vec{x},\tau)=-\frac{da}{d\tau}\frac{\vec{v}(\vec{x},\tau)}{a}-\vec{\nabla}\Phi \eql{eulereq}\\
&&\nabla^2 \Phi = 4\pi G a^2 \bar{\rho}_{\rm m} \delta(\vec{x},\tau) .\eql{fluid3}
\eea
These equations fully determine the time evolution of the local density contrast $\delta$ and the peculiar velocity field $\vec{v}=d\vec{x}/d\tau$.
We can solve these equations perturbatively in Fourier space~\cite{Bernardeau:2002rev}:
\bea
{\delta}(\vk,\tau) &=&\sum_{n=1}^{\infty} {\delta}^{(n)}(\vk,\tau)=\sum_{n=1}^{\infty}D_1^{n}(\tau)\delta_{n}(\vk)  \eql{pert1}\\
{\theta}(\vk,\tau)&=&\sum_{n=1}^{\infty}{\theta}^{(n)}(\vk,\tau) \vs
&=&-\frac{d\,\ln D_1(\tau)}{d\tau}\sum_{n=1}^{\infty}D_1^{n}(\tau)\theta_{n}(\vk) \eql{pert2}
\eea
where $D_1$ is the linear growth factor. 
The first order term ${\delta}^{(1)}$ corresponds to linear evolution. The linear power spectrum is given by this first order term via:
\be 
\langle {\delta^{(1)}}(\vk,\tau){\delta^{(1)}}(\vk',\tau) \rangle =(2\pi)^3 \delta_{\rm D}(\vk+\vk')P_{\rm lin}(k,\tau) \eql{lin}
\ee 
Substituting the perturbative series \ec{pert1} and \ec{pert2} into the Fourier transformed fluid equations \ec{fluid1}-\ec{fluid3} leads to an expression for the second-order density contrast:
\be
{\delta}^{(2)}(\vk,\tau)=\int \frac{d^{3}\vk_{1}}{(2\pi)^3} F_2(\vk_1,\vk-\vk_1){\delta}^{(1)}(\vk_1,\tau) {\delta}^{(1)}(\vk-\vk_1,\tau) \eql{sorder} 
\ee
with 
\be
F_{2}(\vk_1,\vk_2)=\frac{5}{7}+\frac{2}{7}\frac{(\vk_1\cdot \vk_2)^2}{k_1^2 k_2^2}+\frac{\vk_1\cdot \vk_2}{2k_1k_2}(\frac{k_1}{k_2}+\frac{k_2}{k_1}).\eql{f2}
\ee
Note that \ec{pert1}, \ec{pert2} and \ec{f2} are completely accurate only in an Einstein-de Sitter universe. Nonetheless for related calculations in a $\Lambda$CDM universe, the difference is found to be negligible \cite{Takahashi:2008to}, and thus we use the expressions from \ec{pert1} and \ec{f2} throughout this work.\\

Using this expression for $\delta^{(2)}$, we can calculate the two-point correlation of two short-wavelength modes $\vk_s$ and $\vk_s'$, in the squeezed limit $\vk_l=\vk_s+\vk_s'$ with $\vk_s,\vk_s' \gg \vk_l$. Here $\vk_l$ corresponds to a long-wavelength mode, and we suppress the time dependence. To second order,
\bea 
 \langle {\delta}(\vec{k}_s){\delta}(\vec{k}_s') \rangle|_{\vk_s+\vk_s'=\vk_l}&=&
  \langle {\delta}^{(1)}(\vec{k}_s){\delta}^{(2)}(\vec{k}_s') \rangle\vs
  &+&\langle {\delta}^{(2)}(\vec{k}_s){\delta}^{(1)}(\vec{k}_s') \rangle.
\eea 
Substituting \ec{sorder} into the first bracket we get:
\bea 
\langle {\delta}^{(1)}(\vec{k}_s){\delta}^{(2)}(\vec{k}_s') \rangle =  \int \frac{d^3\vec{k}}{(2\pi)^3} F_2 (\vec{k},\vec{k}_s'-\vec{k})\vs
\times \langle {\delta}^{(1)}(\vec{k}_s){\delta}^{(1)}(\vec{k}_s'-\vec{k}){\delta}^{(1)}(\vec{k}) \rangle \eql{integral}
\eea 
When one of the wavenumbers in the 3-point function in \ec{integral} is very small, that mode can be considered as a background mode. The small scale modes evolve in the presence of whatever long wavelength modes happen to be present. Therefore, 
\bea
\langle
\delta^{(1)}(\vk_{s}) \delta^{(1)}(\vk_{s}-\vk){\delta^{(1)}}(\vk) 
\rangle
&=&\langle {\delta^{(1)}}(\vk_{s}) {\delta^{(1)}}(\vk_{s}-\vk) \rangle {\delta^{(1)}}(\vk)  \vs
&+&\langle {\delta^{(1)}}(\vk_{s}) {\delta^{(1)}}(\vk) \rangle {\delta^{(1)}}(\vk_s'-\vk) .
\vs
\eql{contraction}
\eea
The first term on the right occurs when $\vk$ is small and the second when $\vk_s'-\vk$ is small.
Using \ec{lin}, \ec{integral} then becomes:
\bea 
&&\int \frac{d^3\vec{k}}{(2\pi)^3} F_2 (\vec{k},\vec{k}_s'-\vec{k})\langle {\delta}^{(1)}(\vec{k}_s){\delta}^{(1)}(\vec{k}_s'-\vec{k}){\delta}^{(1)}(\vec{k}) \rangle \vs
&=&\int d^3\vk F_2(\vec{k},\vec{k}_s'-\vec{k}) \delta_{\rm D}(\vk_s+\vk_s'-\vk)P_{\rm lin}(k_s){\delta}^{(1)}(\vec{k}) \vs
&+&\int d^3\vk F_2(\vec{k},\vec{k}_s'-\vec{k}) \delta_{\rm D}(\vk_s+\vk)P_{\rm lin}(k_s){\delta}^{(1)}(\vk_s'-\vk) \vs
&=&2F_2(-\vk_s,\vk_s+\vk_s')P_{\rm lin}(k_s){\delta}^{(1)}(\vk_s+\vk_s')
\eea 
where we take advantage of the fact that $F_2$ is a symmetric function. Finally we have:
\be 
\langle {\delta}(\vec{k}_s){\delta}(\vec{k}_s') \rangle =f(\vec{k}_s,\vec{k}_s'){\delta}^{(1)}(\vec{k}_l) \eql{2pt}
\ee 
with
\bea
f(\vec{k}_s,\vec{k}_s')&=&2F_2(-\vec{k}_s,\vec{k}_s+\vec{k}_s')P_{\rm lin}(k_s)\vs
&+&2F_2(-\vec{k}_s',\vec{k}_s+\vec{k}_s')P_{\rm lin}(k_s')       
\eea 
\ec{2pt} suggests that we can estimate long-wavelength modes using short-wavelength modes.\\

\noindent{\it Quadratic Estimator.}\label{sec2}
We can now construct the quadratic estimator for long-wavelength modes starting from \ec{2pt} and summing over as many pairs as possible with weights that maximize the signal to noise. As with the case of CMB lensing, we can write the general form of the estimator by averaging over pairs of short-wavelength modes:
\begin{eqnarray}
\hat{\delta}^{(1)}(\vec{k}_l)=A(\vec{k}_l)\int \frac{d^3 \vec{k}_s}{(2\pi)^3} g(\vec{k}_s,\vec{k}_s'){\delta}(\vec{k}_s){\delta}(\vec{k}_s') \eql{quadest},
\end{eqnarray} 
with $g$ being a weighting function, $\vk_s'=\vk_l-\vk_s$ and $A$  defined by requiring that $\langle \hat{\delta}^{(1)}(\vec{k}_l) \rangle={\delta}^{(1)}(\vec{k}_l)$:
\begin{eqnarray}
A(\vec{k}_l)=\bigg[\int \frac{d^3 \vec{k}_s}{(2\pi)^3} g(\vec{k}_s,\vec{k}_s')f(\vec{k}_s,\vec{k}_s')  \bigg]^{-1} \eql{a}
\end{eqnarray}
In the absence of shot noise the Gaussian noise is given by:
\be 
\langle \hat{\delta}^{(1)}(\vk_{l})\hat{\delta}^{(1)}(\vk_{l}') \rangle = (2\pi)^3 \delta_{\rm D}(\vk_{l}-\vk_{l}')[P_{\rm lin}(k_{l})+N(\vk_l)]
\ee 
with 
\begin{eqnarray}
&&N(\vec{k}_{l})=2A^2(\vk_{l})\vs
&&\times\int \frac{d^3 \vec{k}_{s}}{(2\pi)^3} g^2(\vec{k}_{s},\vk_l-\vec{k}_{s})P_{\rm nl}(k_{s})P_{\rm nl}(|\vk_l-\vk_s|)
\end{eqnarray}
where $P_{\rm nl}$ is the nonlinear power spectrum. Minimizing the noise term we can fix the form of $g$ to be:
\begin{eqnarray}
g(\vec{k}_{s},\vec{k}_{s}')=\frac{f(\vec{k}_{s},\vec{k}_{s}')}{2P_{\rm nl}(k_{s})P_{\rm nl}(k_{s}')}
\end{eqnarray} 
The noise term reduces simply to $N(\vk_l)=A(\vk_l)$. We find by testing that the value of $N$ is very insensitive to the choice of the lower limit of the integration \ec{a}, since most of the contribution comes from large $k_s$.

Assuming Gaussian noise, the projected detectability of a $P(k_l)$ measurement using the quadratic estimator can be expressed as:
\be
\frac{1}{\sigma^{2}(k_l)}=\frac{V k_l^2 \Delta k }{(2\pi)^2}\bigg[\frac{P_{\rm lin}(k_l)}{P_{\rm lin}(k_l)+N(k_l) }\bigg]^2 \eql{error},
\ee
where $V$ is the volume of a survey and we compute the detectability for a set of narrow $k_l$-bins each separated by width $\Delta k$.
In Fig.~\rf{SN}, we show the projected errors on the long-wavelength power spectrum using this quadratic estimator in a large survey. The upper limit of the $\vk_s$ integration in \ec{a} is set to be $0.22\,h\,\rm Mpc^{-1}$. We will see that this choice of the upper limit is reasonable for our current construction at $z=0$. Also notice that $P_{\rm lin}(k_l)$ dominates over $N(k_l)$ in \ec{error} for this upper limit, thus the projected error bars are only slightly wider than the cosmic variance error bars ($N=0$).

\Sfig{SN}{Long-wavelength power spectrum and its error from \ec{error} which can be expressed as $P(k_l)\sigma(k_l)$. We assume a toy survey of boxsize $L=2.5\, h^{-1}\rm\, Gpc$, thus volume $V=L^3$ and width $\Delta k = 2\pi/L$. Moreover, we set the integration range for $\vk_s$ from $0.03 \,h \,\rm Mpc^{-1}$ to $0.22\,h\,\rm Mpc^{-1}$.}

@@

We use the code nbodykit \cite{Hand:2018nby} to measure the Fourier density modes, and \ec{quadest} to estimate the long wavelength modes  from the measured short wavelength modes. How well the estimator works can be seen from Fig.~\rf{hist}, where we show histograms of the ratio of the estimated mode amplitudes $\hat{\delta}(\vk_l)$ to their true amplitudes $\delta(\vk_l)$ for different values of $\vk_l$. The two panels show the differences between a short wavelength mode cutoff of $k_s=0.22\, h\, \rm Mpc^{-1}$ and  $k_s=0.37\,h\, \rm Mpc^{-1}$. Notice that second-order SPT becomes less accurate as shorter wavelengths are used and will produce a bias of our quadratic estimator. The figure shows that individual mode amplitudes are  unbiased when $k_{s,\rm max}= 0.22\,h\, \rm Mpc^{-1}$, while for $k_{s,\rm max}=0.37\,h\, \rm Mpc^{-1}$, the results are biased (the center of the ratio is $\sim 20 \%$ too high).
\Sfig{hist}{Comparison of measured and predicted Fourier mode amplitudes: histograms of number counts of $|\hat{\delta}(\vk_l)/\delta(\vk_l)|$ for two different $k_s$ integration ranges.}

@@

The bottom panel shows the difference between the estimated and true density fields. It is apparent that the differences are much smaller than the overdensities; equivalently the estimator does an excellent job of extracting the large scale density field.\\

\noindent{\it Conclusion.}\label{sec4}
In this paper, we have proposed a new and potentially powerful method to measure long-wavelength modes without having to actually measure large scale structure directly. The estimator works well on an N-Body simulation, so applying this estimator to survey data is an exciting next step. Among the issues that must be faced when dealing with a spectroscopic galaxy survey are: galaxy bias, redshift space distortions \cite{Kaiser:1987rsd}, and light-cone effects. We do not expect any of these to be show-stoppers, so it is tempting to speculate about the possibilities that will open up with this estimate of the large scale density field. 
First, we can hope to measure 3D clustering on scales larger than the matter radiation equality turnover without worrying about large-scale systematic effects. General relativistic effects are strongest on large scales (e.g., \cite{Jeong:2012ls}), and these could be detected. There is evidence of large scale anomalies, in the CMB, that could be confronted with maps of large scale structure obtained with this estimator. Primordial non-Gaussianity generated by inflation leaves an imprint on the largest scales. There is even the possibility of cross-correlating the large-scale matter field with the CMB itself to extract information about the longest wavelength modes in the universe. Since the current epoch of acceleration is a large-scale, late-time effect, there is the possibility of learning about the mechanism responsible for acceleration. Although, as mentioned above, challenges remain, there is also the possibility of using even smaller wavelength modes in our estimator by going to higher order in perturbation theory.\\
